Transformation of forecasts for evaluating predictive performance in an epidemiological context

### 2023-06-20 reply_to_reviews.tex

This illustrates that a quadratic mean-variance relationship can arise from a simple yet not completely unreasonable renewal equation model.

% to model epidemiological processes. We agree with the reviewer that a size parameter of 0.1 is not realistic, but still think that the example is valuable as an illustration.

@@

\blue
Figure 3: is a geometric distribution realistic for epidemiological surveillance data?

\black
We agree that this is not a typical example we would expect to see in an epidemiological setting. It could happen in principle, but only in rather constructed cases. We still believe the example is valuable to illustrate the general point that the rankings between two forecasts can change even for a single observation. We added the two following sentences:

\indented
We note that the chosen example involving a geometric forecast distribution is somewhat constructed; for more similarly shaped pairs of forecast distributions this phenomenon will not typically occur.

\notindented

@@

\blue
L236: “better able to illustrate the effects” — Why does this removal help illustrate the effects? Is such a removal procedure required to be able to apply this transformation approach at all? And if so, what are the requirements/considerations for such a removal procedure?

\black
We updated the paragraph as follows:

\indented
In addition, we filtered out erroneous forecasts that were in extremely poor agreement with the observed data, as defined by any of the conditions listed in Table \ref{tab:erroneous}. Figure \ref{fig:erroneous-forecasts.png} shows the percentage of forecasts removed for each model. Those few (less than 0.2\%  of forecasts for each model) erroneous outlier forecasts had excessive influence on average scores and relative skill scores in a way that was not representative of normal model behaviour. We removed them here in order to better illustrate the effects of the log-transformation on scores that one would expect in a well-behaved scenario. In a regular forecast evaluation such erroneous forecasts should usually not be removed and would count towards overall model scores.

\notindented

\red
L244: “pairwise comparisons” — The necessity of using an approach to handle missingness when combining evaluations should be noted more prominently from the beginning. The same applies to the outlier removal (if necessary) and a value selection.

\black
We added the following short paragraph to Section 2.4 to ensure these aspects are already mentioned in the conceptual part of the paper.

\indented
We note that in the practical evaluation of operational forecasting systems  several additional challenges arise, which we do not study in detail. These concern e.g., the removal of outlying observations and forecasts and the handling of missing forecasts. The solutions we employed in practice are provided in Section 3.1.

\notindented
% I'm not sure where we should note this. This is already more or less the beginning of the example part. 
% And for the theoretical part it's not that relevant, because missingness is somewhat independent of whether we transform forecasts.

\blue
Figure 4: The data (not) selected seems to obscure the key period of predicting the start of a case wave, as we expect the CRPS of log data to emphasize missed case increase predictions there more heavily than the natural scale. Deaths are not a substitute because they have access to a leading indicator and the same sort of misses may not be observed. The evaluations here may also miss some overshooting of the case peak. Which rules from Table SI.2 are coming into play and why? Is there a more complete alternative example?

\black
The data that was removed here was largely due to data revisions, not due to forecasts removed by the filtering on our part. 
We added the percentages of observed values in the main text and added the following two Figures in the SI to make that clearer.

@@

\blue
L24 “is only one” — Seems a bit too similar to “the only one”, “the only”, etc.

\black
We hope that the remainder makes it sufficiently clear that other transformations are possible.

\blue
L46 “WIS is an approximation of the CRPS” — Is it an approximation of CRPS itself or some (2x?) scaling?

\black
The weighted interval score is a weighted sum of interval scores for the individual prediction intervals. When weighing the individual interval scores for the central $1 - \alpha$ prediction intervals by 
$w = \alpha / 2$ (which is generally standard), then the WIS converges exactly to the CRPS for an increasing number of equally spaced prediction intervals. For an illustration see Figure 2 in \cite{bracherEvaluatingEpidemicForecasts2021}.

@@

\textit{The reproduction number $R$ describes the number of people each infected person is expected to infect in turn (irrespective of the time between infections) and therefore is a measure of the strength of epidemic growth. The growth rate $r$ describes the speed of epidemic growth.}

\blue
L60: “immunity” $\xrightarrow{}$ “population immunity”?

\black
We changed this to "population immunity" as suggested.

\blue
L69: This might benefit from a short justification (e.g.,
$\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}$)

@@

\black
This is a great suggestion, thank you. We updated the corresponding paragraph, which now reads:

\indented
If case numbers are evolving based on an exponential process and the  modelling task revolves around estimating and forecasting the reproduction number or the corresponding growth rate, then evaluating forecasts based on the absolute distance between forecast and observed value penalises underprediction (of the reproduction number or growth rate) less than overprediction by the same amount. Consider a forecast that misses the correct growth rate by either $+\epsilon$ or $-\epsilon$:

@@

\begin{equation}
\left| \frac{x_t e^{(r_t + \epsilon) \Delta t} - x_t e^{r_t \Delta t}}{x_t e^{(r_t - \epsilon) \Delta t} - x_t e^{r_t \Delta t}} \right| = e^{\epsilon\Delta t}
\end{equation}

For exponential processes errors on the observed value grow exponentially with the error on the estimated reproduction number or growth rate.

\notindented
\blue
L70: Clarify: is this suggesting to make the forecasting targets themselves the result of applying some standardized growth rate estimator to current+future observations around the desired time period?

@@

\black 
Here we aren't suggesting a specific transformation yet, but rather just say that it would be desirable to score a forecast of the growth rate (regardless of how that can be achieved). Specific suggestions for how to achieve this (either by taking a log transformation or by using another transformation to obtain growth rates directly as mentioned in the Discussion) are discussed in more detail in the remainder of the text. We added the word 'conceptually' to the sentence to make this a bit clearer.

\blue
L99: How are atoms at $\exp x = 0$ handled? How are observations of $y = 0$ handled?

\black
In the theoretical discussion in Section 2 we defined prediction targets to have strictly positive support, so these cases are excluded. In the later parts we follow the common strategy to add an offset before taking the logarithm.

L127: “approximation of the absolute percentage error” — Raises a few questions:
\begin{itemize}
    \item If a relative error metric is desired, why not just divide CRPS/WIS by y then? (Propriety?)\\
    \black
Yes, dividing by the observed value generally disturbs propriety. This approach would correspond to weighting CRPS scores with weights depending on the observed value, and it has been shown that the resulting average scores are no longer proper; see Lerch et al (2017, Statistical Science, preprint: \url{https://arxiv.org/abs/1512.09244})
    \item \blue How does using a (shifted) log transform differ from dividing by (a constant plus) (a mean of) the most recent observation(s)?\\
    \black Like in the related arguments in the paper we address this question for the simpler case of the aboslute error rather than CRPS. Using a shifted log transform corresponds to (in lsight variation of equation (8))
    \begin{equation}
    |\varepsilon^*_t| = |\log (\hat{y_t} + a) - \log(y_t + a)| = \left|\log\left(\frac{\hat{y}_t + a}{y_t + a}\right) \right| \ \ \stackrel{\text{if } \hat{y}_t \ \approx \ y_t}{\approx} \ \ \left| \frac{\hat{y}_t + a}{y_t + a} - 1 \right| \ \ = \ \ \left| \frac{\hat{y}_t - y_t}{y_t + a} \right|.
    \end{equation}
    Dividing by the last observation plus a constant $a$, on the other hand, corresponds to
    \begin{equation}
    |\varepsilon^*_t| = \left|\frac{\hat{y}_t + a}{y_{t - 1} + a} - \frac{y_t + a}{y_{t - 1} + a} \right| = \ \ \left| \frac{\hat{y}_{t - 1} - y_t}{y_{t - 1} + a} \right|.
    \end{equation}
    If we interpret the denominator as an (inverse) weight for the different absolute errors, log transformation thus corresponds approximately equivalent to weighting with the inverse new observation $y_t$ (though better behaved in terms of the resulting incentives). Dividing by the previous value corresponds to weighting with the inverse preceding observation. As autocorrelation is typically strong in the time series we consider, this will yield similar, but not identical results.\\
    In terms of the epidemiological  interpretation, dividing by the last observed value can be interpreted as a score for a multiplicative, rather than an exponential growth rate. % We haven't investigated in detail how the two differ in their behaviour, but would expect them to yield mostly similar results. 
    \item \blue Is this desired? Though already discussed, seems like the major point of doubt here: are we at risk of overemphasizing very-low-activity periods and very-low-population states where standard-deviation-to- mean ratios are much higher?\\
    \black We agree that it may be an issue in practice that low-activity-periods could get overemphasised and therefore believe that an evaluation should show both scores on the natural scale as well as on the log scale. 
\end{itemize}

@@

\red

L127: As APE, RE, and SAPE are defined differently in other sources (e.g., with scaling factor in APE and SAPE, and with y in the denominator and no absolute value in RE), it may be helpful to also indicate sources for all three (e.g., that the RE definition is also in Gneiting, 2011).

\black
We added a reference to Gneiting (2011) for the relative error. SAPE is not defined in that paper, and we reference the paper by Flores (1986) instead. It seems that the earliest occurrence of the SAPE is due to a book by Armstrong (\textit{Long-range Forecasting: From Crystal Ball to Computer}, 1985); however, as we were unable to access the full text of this book we consider Flores (1986) the more helpful reference.

% Is there a good source for the SAPE? It's not in the Gneiting paper. I found this, but not entirely convinced: https://bmcinfectdis.biomedcentral.com/articles/10.1186/s12879-017-2365-1. Wikipedia (\url{https://en.wikipedia.org/wiki/Symmetric_mean_absolute_percentage_error})also suggests Armstrong and Flores.

@@

L146: “$\Bar{\hat{r}}$” — Should be $\hat{\Bar{r}}$, or consider $\hat{\Bar{r}}_t$.

\black
We applied the suggested changes.

\red
L158 “delta method” — Delta method would seem to also require the distribution to be somewhat tightly distributed, concentrated within a nearly-linear region of the transformation function. But putting significant mass on a wider range of values would also mean having non-negligible mass on negative values if “approximately normal”; it would not make sense to apply the logarithm or sqrt transformations in these cases. So perhaps the “tightly distributed” assumption can be taken. But this seems a bit nonobvious and a justification/application of the delta method would be helpful here.

This argument is a common justification for the choice of variance-stabilizing transformations, see e.g., the below excerpt from \textit{Generalized Linear Models With Examples in R} by Dunn and Smyth:
\begin{quote}
Suppose that $y$ has a mean-variance relationship defined by the function $V(\mu)$, with $\text{var}(y) = \phi V(\mu)$. Then, consider a transformation $y^* = h(y) \approx h(\mu) + h'(\mu)(y - \mu)$, from which it can be inferred that
$$
\text{var}(y^*) = \text{var}[h(y)] \approx h'(\mu)^2\text{var}(y).
$$
Hence, the transformation $y^* = h(y)$ will approximately stabilize the variance if $h'(\mu)$ is proportional to $\text{var}(y)^{-1/2} = V(\mu)^{-1/2}$. When $V(\mu) = \mu^2$ [...], the variance-stabilizing transformation is the logarithm, because then $h'(\mu) = 1/\mu$.  When $V(\mu) = \mu$, the variance-stabilizing transformation is the square root because $h'(\mu) = 1/\mu^{1/2}$.
\end{quote}

We added a reference to this source. We nonetheless agree with the referee that this approximation will work best if the distribution is tight and added a short remark to this regard.

The current paragraph reads:

\indented

If the mean-variance relationship is quadratic with $\sigma^2 = c \times \mu^2$, the natural logarithm can serve as the VST \citep{guerreroTimeseriesAnalysisSupported1993}. Denoting by $F_{\log}$ the predictive distribution for $\log(Y)$, we can use the delta method (a first-order Taylor approximation, see e.g., \citealt{Dunn2018}), to show that

\begin{equation}
\mathbb{E}_F[\text{CRPS}\{F_{\log}, \log(y)\}] \approx \frac{\sigma/\mu}{\sqrt{\pi}} 
= \frac{\sqrt{c}}{\sqrt{\pi}}
.
\end{equation}

The addition on ``tightly distributed'' probability mass reads

\indented
We note that while standard in the derivation of variance-stabilizing transformations, the application of the delta method in equations \eqref{eq:taylor_quadratic} and \eqref{eq:taylor_linear} requires the probability mass of $F$ to be tightly distributed. If this is not the case, the approximation may be less accurate.

\notindented

\blue
L159 — Need to discuss / spell out the point of the above calculation (— that it is the same for all $\mu$, when paired with the same $c$?).

\black
We added a sentence to make our point clearer:

\indented
As $\sigma$ and $\mu$ are linked through the quadratic mean-variance relationship ($\sigma = \sqrt{c} \times \mu$), the expected CRPS stays constant regardless of the expeccted value of the data-generating distribution $\mu$.

\notindented

@@

\notindented

\blue
L168: “grows with the variance” — May read like this a linear growth. Consider “grows with the standard deviation”.

\black
We have modified the sentence, which now reads:

\indented
We see that when applying the CRPS on the natural scale, the expected CRPS grows monotonically as the variance of the predictive distribution (which is equal to the data-generating distribution for the ideal forecaster) increases.

\notindented
As the rest of the paragraph talks about the variance, we kept the word variance here as well.

@@

\blue

L180: “data-generating” — Consider “fitted”/“believed”.

\black
In this instance "data-generating" is correct, as this is about a general property of proper scoring rules. We changed the sentence slightly to make this clearer:

\indented
We computed expected CRPS values  for three different distributions, assuming an ideal forecaster with predictive distribution equal to the true underlying (data-generating) distribution.

\notindented
\blue
L186: “incentivizes overconfident predictions” — Needs justifica- tion/citation/removal. This single example disproves propriety but doesn’t establish overconfidence as being encouraged in general.

\black
You are right. We removed that sentences as we do not have an analytical result on whether this holds in general.

\blue
Figure 2: Were any zeros encountered for the negative binomial or Poisson distributions? Was a shifted log transformation used rather than an unshifted log transformation?

@@

\black
Thank you for the suggestion, we edited the text accordingly.

\blue
Figure 5 caption:
\begin{itemize}
    \item “logarithmic” — With what a values? (THe first link to Figure 5 is from before this information was stated.)
    \item “D: [. . . ] all individual scores” — of EuroCOVIDhub-ensemble or any of the 7 included systems?
    \item “relative change” — of what change in scores relative to what quantity?
\end{itemize}

\black
We have updated the caption, which now reads:

\indented
Scores for two-week-ahead forecasts from the EuroCOVIDhub-ensemble (averaged across all forecast dates) for different locations, evaluated on the natural scale as well as after transforming counts by adding one and applying the natural logarithm. D: Corresponding boxplots of all individual scores of the EuroCOVIDhub-ensemble for two-week-ahead predictions. E: Boxplots for the relative change of scores for the EuroCOVIDhub-ensemble across forecast horizons. For any given forecast date and location, forecasts were made for four different forecast horizons, resulting in four scores. All scores were divided by the score for forecast horizon one. To enhance interpretability, the range of visible relative changes in scores (relative to horizon = 1) was restricted to [0.1, 10].

\notindented
\red

@@

L272: “with the standard deviation” — Also note here that this is for approximately normal data distributions?

\black 
We now remind the reader of this assumption by writing

\indented

As (under the simplifying assumption of normality, see Section \ref{sec:methods:vst}) the WIS/CRPS of an ideal forecaster scales with the standard deviation, a value of $\beta = 1$ would imply a quadratic median-variance relationship;

\notindented

@@

\blue
L282: “independent” — Consider “linearly independent”.

\black
We changed the sentence accordingly.

\black

\blue
L288–290: What would stakeholders want this difference to be, and in
which direction? I would guess the sqrt would generally be preferable unless applying some extra transformation after the log to re-emphasize locations/times of interest.

@@

\black
That likely depends on the exact purpose stakeholders have in mind. In many instances it is presumably not desirable to give large weight to smaller forecast targets, and one may therefore e.g. choose to use the sqrt transformation or a different one over the log transformation. On the other hand, in our example the log transformation only gives a little bit more weight to smaller forecast targets, which may be acceptable given that the log transformation has nice properties in terms of its interpretation in an epidemiological context.

\blue
Figure 7 caption: “one score” — Consider clarifying that/whether there is potentially missingness here.

\black
We have updated the Figure caption to make it more easily understandable. It now reads:

\indented
Correlations of rankings on the natural and logarithmic scale. A: Average Spearman rank correlation of scores for individual forecasts. For every individual target (defined by a combination of forecast date, target type, horizon, location), one score was obtained per model. Then, for every forecast target, the Spearman rank correlation was computed between scores on the natural scale and on the log scale for all the models that had made a forecast for that specific target. These individual rank correlations were then averaged across locations and time and are displayed stratified by horizon and target types, representing average accordance of model ranks for a single forecast target on the natural and on the log scale. B: Correlation between relative skill scores. For every forecast horizon and target type, a separate relative skill score was computed per model using pairwise comparisons, which is a measure of performance of a model relative to the others for a given horizon and target type that accounts for missing values. The plot shows the correlation between the relative skill scores on the natural vs. on the log scale, representing accordance of overall model performance as judged by scores on the natural and on the log scale.

\notindented

@@

\blue
Figure 8: What is the meaning of the arrow colors? Initially it seems to match the direction, but there is also blue arrow pointing downward. If it is indicating the direction, please fix this and consider a third color for indicating unchanged ranks.

\black
The colour of the error indicates whether or not relative skill scores improved or deteriorated. We agree this was not at all clear and thank you for pointing this out. We updated the plot legend and added a more detailed explanation to the Figure legend:

\indented
Red arrows indicate that the relative skill score deteriorated when moving from the natural to the log scale, green arrows indicate they improved.

\notindented

\blue
The WIS contribution decomposition suggests that an alternative baseline should be used based on differences in transformed data rather than differences in natural scale data. Or, alternatively, that the a value is too low to give reasonable comparisons (as data that is hard/impossible to distinguish from 0 on a plot on the natural scale corresponds with very high penalization of forecasts considering 0 as a possibility). This could (also) be noted in Figure SI.5’s caption.

\black
We agree that it would be interesting for the Forecast Hubs to explore baseline models based on transformed differences. However, in this application we aimed for coherence with previous analyses of the same data and thus included the baseline models created within the Forecast Hub in real time.

We also agree it is useful to flag the issues surrounding the selection of $a$ more prominently. We added the following sentence to the caption of Figure SI.5:

### 2023-06-20 reply_to_reviews.tex

\indented
One potential way to do deal with this could be to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of $a = 1$.

\notindented

We also added a similar sentence to a paragraph in the main text, which now reads:

\indented

A related issue occurs when the predictive distribution has a large probability mass on zero (or on very small values), as this can translate into an excessively wide forecast in relative terms. This can be seen in Figure SI.5. Here, the dispersion component of the WIS is inflated for scores obtained after applying the natural logarithm because forecasts contained zero in its prediction intervals. To deal with this issue one could potentially choose to use a higher $a$ value when applying a transformation $\log(x + a)$, for example $a = 10$ instead of the $a = 1$ that we chose to use.

\notindented

@@

\blue
L341–342: “natural logarithm as a [. . . ] (VST)” — for what type of relevant distribution(s)?

\blue
L344: “take differences between forecasts on the log scale” — Clarify.

L345: “divide each forecast by the forecast” — Caveat about current forecast formats should be included here / in following sentences, without (one mention of) the idea about directly soliciting forecasts of week-on-week ratios in between.

\black
We updated the relevant paragraph which now reads:

\indented

We suggested using the natural logarithm as a variance-stabilising transformation (VST). This is appropriate for variables that are approximately normally distributed and have a quadratic mean-variance relationship with $\sigma^2 = c \times \mu^2$ (this is e.g. approximately true for the negative binomoial distribution and large $\mu$). Alternatively, the square-root transformation can be appropriate in the case of a Poisson distributed variable \citep{dunnGeneralizedLinearModels2018}. Other VST like the Box-Cox \citep{boxAnalysisTransformations1964} are conceivable as well.
If one is interested in multiplicative, rather than exponential growth rates, one could convert forecasts into forecasts for the multiplicative growth rate by dividing numbers by the last value that was observed at the time the forecast was made. Forecasters would then implicitly predict a separate multiplicative growth rate from today to horizon 1, 2, etc. 
Instead of dividing by the last observed value, another promising transformation would be to divide each forecast by the forecast of the previous week (and analogously for observations), in order to obtain forecasts for week-to-week growth rates. Alternatively, one could also take first differences of values on the log scale. This approach would be akin to evaluating the shape of the predicted trajectory against the shape of the observed trajectory \citep[for a different approach to evaluating the shape of a forecast, see][]{srivastavaShapebasedEvaluationEpidemic2022}. Dividing values by the previous value, unfortunately, is not feasible under the current quantile-based format of the Forecast Hubs, as the growth rate of the $\alpha$-quantile may be different from the $\alpha$-quantile of the growth-rate. However, it may be an interesting approach if predictive samples are available or if quantiles for weekwise growth rates have been collected.

\notindented

@@

\blue
L366: “[...] settings” — (to standardize between measurements, then
could potentially scale by importance to stakeholders)

\black
We slightly adapted the text to make this point clearer:

\indented
It is possible to go beyond choosing a single transformation by constructing composite scores as a weighted sum of scores based on different transformations. This would make it possible to create custom scores and allow forecast consumers to choose and assign explicit weights to different qualities of the forecasts they might care about.

\notindented

@@

\red
Figure SI.2: Would benefit from some more discussion, and perhaps a plot of slightly different quantities. E.g., to select an appropriate value of a, it seems like a common goal would be to deprioritize errors around values of x low enough that we expect there to be primarily noise rather than signal or that are not of concern to policymakers. For count data, this likely includes the values 0 - -10 in any circumstances; how far beyond 10 this range extends depends on the range of realistic growth rates, the forecast horizon, and cleanliness of the data. To assist in this decision, it seems like a similar plot of the evaluations of a particular absolute and/or relative error might be useful, or discussion of how to read it from the current plot. E.g., to illustrate that the same relative error is penalized roughly the same for all the plotted values of a for values of x above 50, but quite differently for $a = 10$ vs. $a = 0.1$ for values of x near 1.

\black
I'm not sure I can follow. In my mind the plot is there to show you that once your values are like 5 times larger than the offset you choose than the offset doesn't have much of an effect. So it's designed to show that for most practical purposes, doing log(x + a) should be mostly equivalent to log(x). So to me it's less about choosing an appropriate a and more about showing that probably it's just fine to go with some random small a. I think Figure SI.7 is more about "choosing an a".

\blue
Figure SI.7:
\begin{itemize}
    \item Also specify sorting variable\&direction on x axis. (Would a point plot vs. the sorting variable be more informative?)
    \item Very informative figure, promote to main text? Also would like to see for different “parts” of an epidemic (e.g., steady low periods, increases, peaks/plateaus, decreases). Deaths figures suggest that balancing scores between locations can be tricky, and both cases and deaths suggests that the investigated transformations can’t do this exactly.
\end{itemize}

\black
We added a statement to the legend "Locations are sorted according to the mean observed value in that location." We also followed your suggestion to make this a main figure.

We would expect that one could find a value for a (e.g. 10?) that would eventually yield a more or less flat line, but we haven't tested this in practice.

@@

Figure 5: “Loaction” → “Location”

Figure SI.7: “Loaction” → “Location”

\black
Thank you very much for pointing these errors out. We have corrected them.

\blue
In several places, e.g., L201, there are links to SI material, when clicked, that instead redirect to main text figures.

\black
We're not sure why this happens and apologise for the confusion. We're also admittedly not sure how to fix this, but expect this will be dealt with during final typesetting.

\blue
\section{Reviewer \#3}

In this manuscript the authors investigate how common scoring rules for probabilistic forecasts can be adjusted by transforming observed data and forecast predictions prior to scoring. They provide three clear motivations for applying such transformations prior to scoring, assess the implications of different transformations in great detail, and evaluate the effects of these transformations using real-world COVID-19 forecasts obtained from the European Forecast Hub. Approaches that can help us improve how we measure the utility of epidemic forecasts for decision-makers are definitely needed. This is a very nice and timely contribution to epidemic forecast evaluation, and I only have a few comments.
